Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-04-03 document/ms.tex
smoothness term should be squared
normaliztion of smoothness prior?
GP prior instead

In the following sections, we will model the occurrence rate
$\rate(\period,\,\radius)$ as a two dimensional grid with $N_\period$ bins in
$\ln\period$ and $N_\radius$ bins in $\ln\radius$.
We choose to fix the bin locations leaving only the bin heights
$\binheights = \{\binheight_{ij}\}_{i=1,j=1}^{N_\period,N_\radius}$ as
free parameters.
In practice, the results seem to be more numerically stable---but the model is
exactly equivalent---when we model the rate as
\begin{eqnarray}
\rate(\period,\,\radius) &=& \rate_0\,p(\period,\,\radius)
\end{eqnarray}
where $\rate_0$ is the total rate of planets in the sample and the relative
probability factor is subject to the constraint
\begin{eqnarray}
1 &=& \int p(\period,\,\radius) \dd\period\dd\radius\quad.
\end{eqnarray}
In this model, the parameter set \ratepars\ includes $\rate_0$ and the heights
of all but one of the bins in $p(\period,\,\radius)$.

Since the choice of bin locations and number is arbitrary, the most
conservative strategy is to choose very small bins---yielding an effectively
non-parametric model---and choose a prior (or regularization) that encodes any
prior expectations about the distribution.
In particular, it seems reasonable to expect the distribution to be
``smooth''.
The most extreme version of this model would be a completely non-parametric
rate function---generated by, for example, a Gaussian process
(\citealt{poiss-gp})---but this discussion is beyond the scope of this
\paper.
Instead, we'll simply choose small bins and add a term to the likelihood
function that penalizes large changes between neighboring bins.
Specifically, we choose a term of the form
\begin{eqnarray}
\ln p(\binheights\,|\,\smoothreg_\period,\,\smoothreg_\radius) &=&
-\frac{1}{2}\sum_{i=2}^{N_\period}
\frac{|\binheights_{i,\cdot} - \binheights_{i-1,\cdot}|}{\smoothreg_\period}
-\frac{1}{2}\sum_{j=2}^{N_\radius}
\frac{|\binheights_{\cdot,j} - \binheights_{\cdot,j-1}|}{\smoothreg_\radius}
-\frac{1}{2}\ln (\smoothreg_\period\,\smoothreg_\radius)
\quad.
\end{eqnarray}
In our final analysis, we'll marginalize over the regularization strengths
$\smoothreg_\period$ and $\smoothreg_\radius$ but we'll also explore the
effects of this choice below.

>>>>

In the following sections, we will model the occurrence rate
$\rate(\period,\,\radius)$ as a two dimensional grid with $N_\period$ bins in
$\ln\period$ and $N_\radius$ bins in $\ln\radius$.
We choose to fix the bin locations leaving only the bin heights
$\binheights = \{\binheight_{ij}\}_{i=1,j=1}^{N_\period,N_\radius}$ as
free parameters.

\dfmtodo{What about the normalization?}

Since the choice of bin locations and number is arbitrary, the most
conservative strategy is to choose very small bins---yielding an effectively
non-parametric model---and choose a prior (or regularization) that encodes any
prior expectations about the distribution.
In particular, it seems reasonable to expect the distribution to be
``smooth''.
The most extreme version of this model would be a completely non-parametric
rate function---generated by, for example, a Gaussian process
(\citealt{poiss-gp})---but this discussion is beyond the scope of this
\paper.
Instead, we'll simply choose small bins and add a Gaussian process prior
(\dfmtodo{CITE}) on the logarithms of the bin heights
\begin{eqnarray}
\ln p(\binheights\,|\,) &=&
-\frac{1}{2}\,\bvec{r}^\mathrm{T}\,\bvec{K}\,\bvec{r}
-\frac{1}{2}\ln \det \bvec{K} - \frac{N_\period\,N_\radius}{2}\ln 2\pi
\end{eqnarray}
where $\bvec{r}$ is the vector of residuals
\begin{eqnarray}
\bvec{r} &=& \left(\begin{array}{cccc}
\binheight_1-\mu & \binheight_2-\mu & \cdots & \binheight_N-\mu
\end{array}\right)^\mathrm{T}
\end{eqnarray}
and $\bvec{K}$ is the kernel matrix generated by
\begin{eqnarray}
K_{ij} &=&
\end{eqnarray}
